Ensemble Methods for Robust Feature Selection

### 2016-08-02 paper/paper.tex
Added analyse about artificial data, and FS distribution

Symmetrical Uncertainty (SU) \citep{press1996numerical} is entropy based measurement. It is the normalized mutual information between the feature and the class. The mutual information measures the influence of the knowledge of the class on the feature value in terms of entropy and normalizes it. More details about mutual information can be found in \cite{paninski2003estimation}.
\begin{equation}
  \label{eq:su}
  SU(F,C) = 2 \frac{H(F) - H(F|C)}{H(F) + H(C)} \textrm{, where } H(\cdot) \textrm{ is the entropy.}
\end{equation}

The second feature selection is RELIEFF \citep{kononenko1997overcoming}. For each sample $x_i$ the k nearest neighboring samples, using the $L_1$ norm, belonging the same class $\{\textrm{Near-hit}_{i, j}\}_{j=1..k}$ and the ones of the opposite class $\{\textrm{Near-miss}_{i, j}\}_{j=1..k}$ are selected. Then the difference is taken to update the weight vector representing how well each feature distinguishes samples of the two classes, as can be seen in equation \ref{eq:relief}. In all the experiments k was set to 5, which, empirically, gives satisfactory results. 
\\
?? \\
In the standard case of RELIEF for comparisment the euclidean distance is used for numerical values. This inherits the
assumption that features which are nearer to each other are also more similar, which should not be correct in each application
area.\\
??
\begin{equation}
  \label{eq:relief}
  W_i = W_i - \frac{1}{k}\sum_{j=1}^{k}|x_i - \textrm{Near-hit}_{i,j}| + \frac{1}{k}\sum_{j=1}^{k} |x_i - \textrm{Near-miss}_{i,j}|
\end{equation}

@@

The worst ranked features are then discarded and the procedure is repeated until only a subset of features is left. In the experiments, $10\%$ of the size of all the remaining features were discarded at each step. This means that. The procedure was repeated until a subset of the size of $1\%$ of all features was left.

The last feature feature selection method used is a RFE with Lasso.

\begin{equation}
  \min_w \quad \frac1{2 \, \textrm{\#features}} \, \|y - Xw\|^2_2 + \alpha \, \|w\|_1
\end{equation}

\section{Ensemble methods}

@@

\begin{description}[align=left]
\item [Linear aggregation :] The average of the weights was taken as the new weight. Two version were used, one where a normalization of the weight sum is applied and the other without it.
\item [Non-linear aggregation :] To emphasize features with importance, the weights are non-linearly transformed with the equation \ref{eq:em_non_linear}. $\arctan$ in in combination with $\exp$ to considerably shrink the low weights. Finally the weights sum is normalized.

\begin{equation}
  \label{eq:em_non_linear}
  W_i = \frac{f \left( W_i \right)}{\sum_k f \left( W_k \right)} \quad with \quad f : x \rightarrow \exp( \arctan(x - 1) )
\end{equation}

\item [Performance related aggregation :] For each feature selector the best 1\% of the features were selected and used to measure the classification performance for multiple classifiers. The feature selection methods result were weighted accordingly to their accuracy. It was used in combination with the precedent ensemble methods.
\end{description}

@@

\subsection{Results}
- Do SVM\_RFE with exponential rising
- SU achieved high values in the paper, in our results SU weighted a lot of features the 
same; therefore for robustness same weights have to be shuffled to not accidental increase
robustness by order of features
- The noise features of the NIPS 2003 challenge data sets are highly correlated with the
real feature; therefore no improvement has been made with feature selection in terms of 
accuracy
% Acknowledgements should go at the end, before appendices and references

@@

\subsubsection{Different combinations of feature selectors}
The presented ensemble methods were used with all possible combinations of feature selectors. For each combination, the average of the results was taken, and the error plot corresponds to the mean of the standard deviation of each ensemble method and each dataset. While the accuracy does not change considerably with the different combinations, the robustness is heavily affected by the presence or not of the feature selector Lasso as can be seen in the figure \ref{fig:combination_robustness}. This is due to the low robustness of the feature selector Lasso itself. On the opposite SU and RELIEFF being quite robust, their combination, with or without SVM, are the most robust.

\begin{figure}
\centering \includegraphics[]{images/Robustness_of_the_different_combinations.pdf}
\caption{Robustness of the different combination of feature selectors (RLF and LSO corresponds respectively to RELIEFF and Lasso)}
\label{fig:combination_robustness}
\end{figure}

@@

\appendix
\section*{Appendix A.}
For the feature selection methods SU and RELIEF e used the scikit-feature library from \cite{Li-etal16}.
\label{app:some appendix}

% Note: in this sample, the section number is hard-coded in. Following
% proper LaTeX conventions, it should properly be coded as a reference:

>>>>

Symmetrical Uncertainty \citep{press1996numerical} is entropy based measurement. It is the normalized mutual information between the feature and the class. The mutual information measures the influence of the knowledge of the class on the feature value in terms of entropy and normalizes it. More details about mutual information can be found in \cite{paninski2003estimation}.
\begin{equation}
  \label{eq:su}
  SU(F,C) = 2 \frac{H(F) - H(F|C)}{H(F) + H(C)} \textrm{, where } H(\cdot) \textrm{ is the entropy.}
\end{equation}

The second feature selection is RELIEFF \citep{kononenko1997overcoming}. For each sample $x_i$ the k nearest neighboring samples, using the $L_1$ norm, belonging the same class $\{\textrm{Near-hit}_{i, j}\}_{j=1..k}$ and the ones of the opposite class $\{\textrm{Near-miss}_{i, j}\}_{j=1..k}$ are selected. Then the difference is taken to update the weight vector representing how well each feature distinguishes samples of the two classes, as can be seen in equation \ref{eq:relief}. In all the experiments k was set to 5, which, empirically, gives satisfactory results. 
\begin{equation}
  \label{eq:relief}
  W_i = W_i - \frac{1}{k}\sum_{j=1}^{k}|x_i - \textrm{Near-hit}_{i,j}| + \frac{1}{k}\sum_{j=1}^{k} |x_i - \textrm{Near-miss}_{i,j}|
\end{equation}

@@

The worst ranked features are then discarded and the procedure is repeated until only a subset of features is left. In the experiments, $10\%$ of the size of all features were discarded at each step. The procedure was repeated until a subset of the size of $1\%$ of all features was left. Since each step $10\%$ of the features are discarded, the overall process takes $10$ steps. Therefore a SVM is $10$ times needed to be evaluated. Another method would be to discard only $10\%$ of the remaining features. A more gradually ascend of feature weights can be achievedthis way. However this was not computable with our ressources. For the hpyerparameter $C$ a grid search on the set $\{1, 10, 100, 1000\}$ was done. The parameters were evaluated with stratisfied 5-fold cross-validation.

The last feature selection method used is Lasso logistc regression \citep{wheeler2010lasso}. It is a logistic regression with a $L_1$ norm as penalty for the weight vector. For the hyperparameter $C$ a grid search with $10$ different parameters in logarithmic scale between $10^{-5}$ and $10^3$ was done. The parameters were evaluated with a stratisfied 3-fold cross-validation.

\begin{equation}
  \min_{w,c} \quad \|w\|_1 + C\sum_{i=1}^n \log(\exp(y_i(x_i^Tw + c))+1)
\end{equation}

The advantage of this method is that we obtain a sparse weight vector as solution. This means only a small subset of features will have a significant high weight.
\section{Ensemble methods}

@@

\begin{description}[align=left]
\item [Linear aggregation :] The average of the weights was taken as the new weight.
\item [Non-linear aggregation :] Two methods were used, first averaging powers of the weights and secondly aggregating linearly the minimum, maximum and the average of the weights.
\item [Performance related aggregation :] For each feature selector the best 1\% of the features were selected and used to measure the classification performance for multiple classifiers. The feature selection methods result were weighted accordingly to their accuracy.
\end{description}

@@

The classification performance is estimated with a cross validation of 10 folds with multiple classifiers, trained and tested on the subset of features. 4 different classifiers are used: Lasso logistic regression, linear SVM, k-nearest neighbor, Random Forest.

\subsection{Results}
- SU achieved high values in the paper, in our results SU weighted a lot of features the 
same; therefore for robustness same weights have to be shuffled to not accidental increase
robustness by order of features

\subsubsection{Properties of FS}
\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{feature_weights_hist_arcene_artificial.png}
  \caption{The histograms of the feature weights for all used feature selectors on the arcene and artificial data set.}
  \label{fig:feature_weights_hist_arcene_artificial}
\end{figure}
In figure \ref{fig:feature_weights_hist_arcene_artificial} the histograms of the features weight can be seen for the arcene and the artificial data set. It is observable that the distribution of the feature weights have similiar shapes on each data set for RELIEFF, SVM RFE, Lasso logistic regression. In addition the distributions are centered around the same area. RELIEFF and Lasso seem to always have a strong peak. Lasso even stronger than RELIEFF, which is an effect of the sparsity of Lasso. On the other hand SU varies a lot on different data sets and does not seem to have a consistent shape. For some data sets like the artifical data set SU mainly outputs one value for all features. On others like arcene it is greatly distributed. The histograms for all data sets can be seen in the appendix in figure \ref{fig:feature_weights_hist}.

\subsubsection{Results to artificial data set}
\begin{figure}
  \centering 
  \begin{subfigure}[b]{0.48\textwidth}
      \includegraphics[width=\textwidth]{feature_weights_plot_artificial_only_features.png}
      \caption{Features weights for the real features.}
      \label{fig:feature_weights_plot_artificial_only_features}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{feature_weights_plot_artificial_all.png}
      \caption{Features weights for all features.}
      \label{fig:feature_weights_plot_artificial_all}
  \end{subfigure}

  \caption{The features weights for once for only the real features and once for all. In figure \ref{fig:feature_weights_plot_artificial_only_features} in the green area are the real multivariate features and in the cyan area are the real univariate features. In figure \ref{fig:feature_weights_plot_artificial_all} in the green are all real features, in the red area are the fake multivariate features and in the orange area are the fake univariate features. }
  \label{fig:feature_weights_plot_artificial}
\end{figure}
\begin{figure}[h!]
  \centering
  \caption{The histograms of the feature weights for all used feature selectors on the arcene and artificial data set.}
\end{figure}
The weights of the features can be seen in figure \ref{fig:feature_weights_plot_artificial}.
All feature selectors achieved similar results for the artificial data set except of SU. SU does not rank iall features with a zero weight which makes SU fully random on this data set. On the contrary the other feature selectors recognize parts of the inherent structure of the data. The univariate features are in general ranked lower than the multivariate features. There seems to be no diffierence in weighting between the real und fake univariate features. However the real multivariate features are ranked significantly higher than the fake multivariate features. As a consequence only multivariate features are selected. The reason why the feature selectors weight the real univariate features this low is that the univariate features do not contribute significantly to the classification. If we use only the real multivariate features with the same classification function with which the labels were created, we achieve an accuracy of $0.98$. On the other hand using the real univariate features we only achieve an accuracy of $0.52$. Therefore a feature which correlates with other significant features has significantly more impact on the classification of a data point than a uniform feature.
Furthermore the feature seletors were evaluated after how many real features they capture in the $100$ the highest weighted features. RELIEFF achieved the highest precision of $0.5$, SVM RFE achieved $0.48$, then Lasso with $0.24$ and SU with $0$.

@@

@@

\appendix
\section*{Appendix A.}
For the feature selection methods SU and RELIEFF we used the scikit-feature library from \cite{Li-etal16}.
For SVM, RFE, Lasso logistic regression we used the scikit-sklearn library from \cite{scikit-learn}
\label{app:appendixa}
\section*{Appendix B.}
\label{app:appendixb}

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{feature_weights_hist.png}
  \caption{The histogram of the feature weights for all feature selectors used for all data sets used.}
  \label{fig:feature_weights_hist}
\end{figure}
% Note: in this sample, the section number is hard-coded in. Following
% proper LaTeX conventions, it should properly be coded as a reference:
